Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation

### 2020-12-08 xrb2/paper.tex
more intro

Considerable evidence suggests that ignition in an X-ray burst starts
in a localized region and then spreads across the surface of the
neutron star~\citep{bhattacharyya:2007,chakraborty:2014}.  We continue
our study of flame spreading through fully compressible hydrodynamics
simulations of the flame.  Building on our previous
paper~\citep{flame_wave1}, we relax the approximations we made
previously (artificially boosting the speed of the flame in order to
reduce the computational cost) and explore how the flame properties
depend on rotation rate and the thermal structure of the neutron
star. This new set of realistic simulations is possible because of the
work done to offload our simulation code, \castro~\cite{castro_joss}, to GPUs, where it
runs significantly faster.

We investigate the effect of rotation rate on the flame. With the exception of (citation to the weird 11 Hz flame \citep{altamirano2010discovery}), most observations of XRBs come from sources with rotation rates of 200-600Hz \citep{bilous:2019,galloway:2020} \MarginPar{check whether this is true for XRBs or just burst oscillations}. There are a number of factors that could explain this lack of observations below 200 Hz. It could be that there is some physical process which inhibits the flame ignition and/or spread at lower rotation rates. It could be that burst oscillations at lower rotation rates are smaller in amplitude and therefore more difficult to detect. It could be that it's not got anything to do with the flame at all, but that neutron stars in accreting low mass X-ray binaries rarely have rotation rates below 200 Hz \MarginPar{look into this?}.

@@

Previous work has predicted that crustal heating is stronger a lower accretion rates and weaker at higher accretion rates, due to the effect of neutrino losses \citep{Cumming2006,johnston:2019}. In our models, we keep the crust at a constant temperature, so by increasing this temperature we can effectively increase the crustal heating and mimic the effects of lower accretion rates. A shallow heating mechanism of as yet unknown origin has been found necessary to reproduce observed properties of XRBs \citep{Deibel2015,Turlione2015,Keek2017}.

>>>>

Considerable evidence suggests that ignition in an X-ray burst starts
in a localized region and then spreads across the surface of the
neutron star~\citep{bhattacharyya:2007,chakraborty:2014}.  We continue
our study of flame spreading through fully compressible hydrodynamics
simulations of the flame.  Building on our previous
paper~\citep{flame_wave1}, we relax the approximations we made
previously (artificially boosting the speed of the flame in order to
reduce the computational cost) and explore how the flame properties
depend on rotation rate and the thermal structure of the neutron
star. This new set of realistic simulations is possible because of the
work done to offload our simulation code, \castro~\citep{castro_joss}, to GPUs, where it
runs significantly faster.

We investigate the effect of rotation rate on the flame. With the
exception of \citet{altamirano2010discovery}, most observations of
XRBs come from sources with rotation rates of 200-600Hz
\citep{bilous:2019,galloway:2020} \MarginPar{check whether this is
  true for XRBs or just burst oscillations}. There are a number of
factors that could explain this lack of observations below 200 Hz. It
could be that there is some physical process which inhibits the flame
ignition and/or spread at lower rotation rates. It could be that burst
oscillations at lower rotation rates are smaller in amplitude and
therefore more difficult to detect. It could be that it's not got
anything to do with the flame at all, but that neutron stars in
accreting low mass X-ray binaries rarely have rotation rates below 200
Hz \MarginPar{look into this?}.

@@

The temperature structure of the accreted fuel layer can also affect
the flame propagation.  \citet{Timmes00} showed that laminar helium flames 
have higher speeds when moving into hotter upstream fuel.
It is known that crustal heating is stronger at lower
accretion rates and weaker at higher accretion rates, due to the
effect of neutrino losses \citep{Cumming2006,johnston:2019}. In our
models, we keep the crust at a constant temperature, so by increasing
this temperature we can effectively increase the crustal heating and
mimic the effects of lower accretion rates. A shallow heating
mechanism of as yet unknown origin has been found necessary to
reproduce observed properties of XRBs
\citep{Deibel2015,Turlione2015,Keek2017}.
